Mutiple level calculation 4

### 2023-04-15 posts/rtables/multiple_levels_calculation_4.qmd
Create multiple_levels_calculation_4.qmd

---
title: 'Mutiple level calculation 3'
date: '2023-04-10'
author: Qing Zou
categories: ['R', 'rtables']
description: 'summary counts by summarize_row_groups'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
format:
  html:
    page-layout: full

---

@@

  if (to_html) {
    flextable::htmltools_value(ft)
  } else {
    ft
  }
}
count_items_cfun <- function(df,
                        labelstr,
                        .spl_context,
                        n_var,
                        n_unique_keys = NULL,
                        N_var = NULL,
                        N_index = NULL,
                        N_unique_keys = NULL,
                        empty_row = FALSE,
                        hide_value = FALSE,
                        format = "n/N(%)") {

@@

## Method 2

```{r cal2}
#| echo: false
#| eval: true
df <- data.frame(SUBJID = c(rep('S1', 5), rep("S2", 4), rep("S3", 2)),
                 ASEQ = c(1:5, 1:4, 1:2),
                 GROUP = factor(c(rep('A', 5), rep("B", 4), rep('A', 2)), levels = c("A", "B")),
                 PARAMCD = c('TEST_1', 'TEST_1', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_2', 'TEST_1', 'TEST_2'),
                 CRIT = c('LEVEL_2_1','LEVEL_2_1', 'LEVEL_2_2', NA_character_, 'LEVEL_2_a', 'LEVEL_2_1', NA_character_, 'LEVEL_2_a', 'LEVEL_2_b', NA_character_, 'LEVEL_2_b'))
```
Another powerful function rtables provide to us is the function `summarize_row_groups`, which can be used to add a content row of summary counts.

This function need a different data structure from the `rtable::analyze()`.
`rtable::analyze()` is kind of the columns-based analysis, the `summarize_row_groups` is more like the rows-based analysis.

To build the data:

- create the 1st level columns by `PARAMCD`, `TEST_1_level_1`, `TEST_2_level_1`
- Create the 2rd level columns by `CRIT`, `TEST_1_level_2`, `TEST_2_level_2`

```{r cal2df}
#| echo: true
#| eval: true

param_coding <- c('TEST_1' = 'This is test 1',
                  'TEST_2' = 'This is test 2')
df <- df %>% mutate(PARAML = param_coding[PARAMCD])
tmp <- df
tmp[, str_c(names(param_coding), '_level_1')] <- NA_character_

# tmp1 <- df %>% mutate(TEST_1_level_1 = ifelse(PARAMCD == 'TEST_1', PARAMCD, NA_character_),
#                       TEST_2_level_1 = ifelse(PARAMCD == 'TEST_2', PARAMCD, NA_character_),
#                       TEST_1_aval_1 = ifelse(!is.na(TEST_1_level_1), 1, NA_character_),
#                       TEST_1_aval_1 = ifelse(!is.na(TEST_2_level_1), 1, NA_character_),
#                       TEST_1_level_2 = ifelse(PARAMCD == 'TEST_1' & !is.na(CRIT), CRIT, NA_character_),
#                       TEST_2_level_2 = ifelse(PARAMCD == 'TEST_2' & !is.na(CRIT), CRIT, NA_character_),
#                       TEST_1_aval_2 = ifelse(!is.na(TEST_1_level_2), 1, NA_character_),
#                       TEST_2_aval_2 = ifelse(!is.na(TEST_2_level_2), 1, NA_character_)
#                       )

tmp1 <- tmp %>%
  mutate(
    across(
      .cols = c("TEST_1_level_1", "TEST_2_level_1"),
      .fns = \(x)  factor(ifelse(grepl(gsub("_level_1", "",cur_column()), PARAMCD), PARAML, x))
    )
    ,
    across(c(do.call(str_c, expand.grid(names(param_coding), c('_level_1')))),
           ~ifelse(!is.na(.x), 1, .x),
           .names = "{gsub('level','aval', .col)}")
    )

tmp1[, str_c(names(param_coding), '_level_2')] <- NA_character_

fmt <- list(
  'TEST_1_level_2' = c('LEVEL_2_1' = '1 of level 2',
               'LEVEL_2_2' = '2 of level 2',
               'LEVEL_2_3' = '3 of level 2'),
  "TEST_2_level_2" = c('LEVEL_2_a' = 'a of level 2',
               'LEVEL_2_b' = 'b of level 2',
               'LEVEL_2_c' = 'c of level 2')
)

tmp2 <- tmp1 %>%
  mutate(
    across(
      .cols = c("TEST_1_level_2", "TEST_2_level_2"),
      .fns = \(x)  factor(ifelse(grepl(gsub("_level_2", "",cur_column()), PARAMCD) & !is.na(CRIT) , fmt[[cur_column()]][CRIT], x), levels = fmt[[cur_column()]])
    ),
    across(c(do.call(str_c, expand.grid(names(param_coding), c('_level_2')))),
           ~ifelse(!is.na(.x), 1, .x),
           .names = "{gsub('level','aval', .col)}")
  )
```

## Wrap the analysis function

```{r a23s}
#| echo: true
#| eval: true
#| message: FALSE

tbl_recipe <- basic_table() %>%
  rtables::split_cols_by(var = "GROUP") %>%
  add_overall_col("All") %>%
  ## 1st-level row-wise splitting, and its column-wise analysis function
  rtables::split_rows_by("TEST_1_level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
  rtables::summarize_row_groups(cfun = count_items_cfun,
                                extra_args = list(n_var = "TEST_1_aval_1", n_unique_keys = "SUBJID", hide_value = TRUE)) %>%
  ## 2nd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
  rtables::split_rows_by("TEST_1_level_2") %>%
  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
    n_var = "TEST_1_aval_2", n_unique_keys = c("SUBJID"), N_var = "TEST_1_aval_1", N_unique_keys = "SUBJID",
    empty_row = TRUE
  ))

col_counts_df <- tmp2 %>%
  distinct(SUBJID, GROUP)
col_counts <- table(col_counts_df$GROUP)
col_counts <- c(col_counts, sum(col_counts))
test1_tbl <- tbl_recipe %>%
  rtables::build_table(tmp2, col_counts = col_counts)
```
```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(test1_tbl, to_html = FALSE)
```

```{r a23ssdf}
#| echo: true
#| eval: true
#| message: FALSE

    # get current value to check if it match the last value of levels, for empty row
    # !Start
    var_levels <- length(.spl_context$split)
    curr_var_level_val <- .spl_context$value[[nrow(.spl_context)]]
    last_var_level_val <-
      .spl_context$full_parent_df[[nrow(.spl_context)]][[.spl_context$split[var_levels]]] %>%
      levels() %>%
      tail(1)
    if (curr_var_level_val == last_var_level_val & empty_row) {
      ret <- in_rows(
        .names = c(labelstr, " "),
        .list = list(c(format_values_helper(n, parent_N, type = format)), NULL),
        .formats = c(sprintf_format_helper(n, parent_N, type = format), "xx")
      )
    } else {
      rc <- rcell(
        x = format_values_helper(n, parent_N, type = format),
        format = sprintf_format_helper(n, parent_N, type = format)
      )
      ret <- in_rows(.names = labelstr, rc)
    }
    # !End
  } else {
    rc <- rcell(
      x = ifelse(hide_value, "", n),
      format = "xx"
    )
    ret <- in_rows(.names = labelstr, rc)
  }
  return(ret)
}

tbl_recipe <- basic_table() %>%
  rtables::split_cols_by(var = "GROUP") %>%
  add_overall_col("All") %>%
  ## 1st-level row-wise splitting, and its column-wise analysis function
  rtables::split_rows_by("TEST_2_level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
  rtables::summarize_row_groups(cfun = count_items_cfun,
                                extra_args = list(n_var = "TEST_2_aval_1", n_unique_keys = "SUBJID", hide_value = TRUE)) %>%
  ## 2nd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
  rtables::split_rows_by("TEST_2_level_2") %>%
  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
    n_var = "TEST_2_aval_2", n_unique_keys = c("SUBJID"), N_var = "TEST_2_aval_1", N_unique_keys = "SUBJID",
    empty_row = TRUE
  ))

test2_tbl <- tbl_recipe %>%
  rtables::build_table(tmp2, col_counts = col_counts)
```
```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(test2_tbl, to_html = FALSE)
```

```{r a23ssdfdf}
#| echo: true
#| eval: true
#| message: FALSE

tbl2 <- rbind(test1_tbl, test2_tbl)

```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(tbl2, to_html = FALSE)
```

>>>>

---
title: 'Mutiple level calculation 4'
date: '2023-04-11'
author: Qing Zou
categories: ['R', 'rtables']
description: 'multiple level counts more than 2 levels'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

@@

  if (to_html) {
    flextable::htmltools_value(ft)
  } else {
    ft
  }
}

@@

## Summary counts more than 2rd level

```{r cal2}
#| echo: false
#| eval: true
df <- data.frame(SUBJID = paste0("S", 1:10),
                 GROUP = factor(sample(c("A", "B", "C"), 10, replace = T), levels = c("A", "B", "C")),
                 AVAL = sample.int(20, 10)
                 )

df <- df %>% mutate(level_1 = ifelse(3 < AVAL & AVAL <=18, "level 1", NA_character_),
                    level_2 = ifelse(5 < AVAL & AVAL <=15, "level 2", NA_character_),
                    level_3 = ifelse(8 < AVAL & AVAL  <=12, "level_3", NA_character_)
                    )
df <- df %>% mutate(across(paste0('level_', 1:3),
                           ~factor(.)))
df <- df %>% mutate(across(paste0('level_', 1:3),
                           ~ifelse(!is.na(.x), 1, .x),
                           .names = "{gsub('level','aval', .col)}"))
```
```{r cal2sd }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

```{r cal2dfg}
#| echo: true
#| eval: true

col_counts <- table(df$GROUP)
col_counts <- c(col_counts, sum(col_counts))

tbl_recipe <- basic_table() %>%
  rtables::split_cols_by(var = "GROUP") %>%
  add_overall_col("All") %>%
  ## 1st-level row-wise splitting, and its column-wise analysis function
  rtables::split_rows_by("level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
  rtables::summarize_row_groups(cfun = count_items_cfun,
                                extra_args = list(n_var = "aval_1", n_unique_keys = "SUBJID", hide_value = FALSE)) %>%
  ## 2nd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
  rtables::split_rows_by("level_2") %>%
  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
    n_var = "aval_2", N_var = "aval_1", N_unique_keys = "SUBJID",
    n_unique_keys = c("SUBJID"), empty_row = FALSE
  )) %>%
  ## 3rd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
  rtables::split_rows_by("level_3") %>%
  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
    n_var = "aval_3", n_unique_keys = c("SUBJID"), N_var = "aval_1", N_unique_keys = "SUBJID", N_index = 1,
    empty_row = TRUE
  ))

tbl_test3 <- tbl_recipe %>%
  rtables::build_table(df, col_counts = col_counts)
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(tbl_test3, to_html = FALSE)
```
